---
title: "Statistic Inference Project Assignment Part1"
author: "Pete Schnettler"
date: "January 13, 2016"
output: html_document
---
# Overview

## Comparing the simulated mean and variance with the theoretical values

We will run 1000 rounds of simulation of 40 exponentials with $\lambda = 0.2$,
using a fixed seed, and comparing the distribution of the simulated mean
and variance with the theoretical value of $1 / \lambda$:


```{r set_global_options, cache=TRUE}
nSim <- 1000
nVals <- 40
lambda <- 0.2
set.seed(456)

#Create exponential distribution and transpose into a matrix 
simMatrix <- t(replicate(nSim, rexp(nVals, lambda)))

#Create a datafrom consisting of the Mean and variance
df <- data.frame(
        Mean=c(mean(rowMeans(simMatrix)), 1/lambda),
        Variance=c(mean(apply(simMatrix, 1, var)), 1/lambda^2))
rownames(df) <- c("Sample", "Theoretical")
library(knitr)
kable(df)
```



# 1.Show the sample mean and compare it to the theoretical mean of the distribution 
The Sample mean is 4.984081 while the Theoretical mean is 5.0 while are within a couple of decimal points, as expected by the Central Limit Theorem.

# 2.Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution
The sample variance 24.7 and theoretical variance 25 are also close

# 3.Show that the distribution is approximately normal

Also, according to the CLT, the distribution of the simulated means 
should be approximately normal. .

```{r}
library(ggplot2)

#Calculate the row means
meanVals <- rowMeans(simMatrix)


zMean <- (meanVals - mean(meanVals)) / sd(meanVals)
qplot(zMean, geom = "blank") +  
  stat_function(fun = dnorm, aes(colour = 'Normal')) +                       
  geom_histogram(aes(y = ..density..), alpha = 0.4, binwidth=.35) +
  geom_vline(xintercept=0, colour="red", linetype="longdash") +
  scale_colour_manual(name = 'Density', values = c('red', 'blue')) +
  ylab("Density") + xlab("z") + 
  ggtitle("Distribution of means of 40 Exponential Distribution Lambda of 0.2") +
  theme_bw() + 
  theme(legend.position = c(0.85, 0.85))
```

The plot of the means above appears approximately normal

